Preprocessing and exploration of the narwhal data for the analysis with smoothSDE

### 2024-04-26 R/preprocessing/data_preprocessing.Rmd
some changes in exploratory graphs

```{=tex}
\begin{itemize}
\item  Drop missing latitude and longitude
\item Keep GPS positions indicated by value 1 in \texttt{pos\_org} column
\item  Project to UTM zone 26. The narwhal positions are in latitude and longitude.To analyze the track and visualize it better we project it on a plane in UTM coordinates.
The positions of the narwhals are in the UTM zone 26 north (in Greenland).
\item Add a column "speed" with empirical horizontal speed obtained from the GPS positions (not interpolated)
\end{itemize}

@@

dataprep=data[!(is.na(data$Lat)) & !(is.na(data$Long)),]
dataprep=dataprep[dataprep$pos_org==1,]
nrows=length(dataprep[,1])

@@

```{r display land , fig.cap="Land polygons"}
plot_land
```

@@

```{r compare distances to shore}
plot(res_land$min_distances/1000-dataprep$DistanceShore)
plot(res_coastline$min_distances/1000-dataprep$DistanceShore)
```

```{r summary distances to shore}
summary(res_land$min_distances)
summary(res_coastline$min_distances)
summary(dataprep$DistanceShore*1000)
```
Coastline distances are a bit closer to original distances in the dataset.
Anyway, we shall keep distance computed from the land in the sequel, since the land polygons will be the reference domain
for constrained movement of the narwhals.

@@

check_plot2=ggplot()+geom_sf(data=land_utm$geometry,fill=NA)+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=df,aes(x,y,col=(DistanceShore<0.1)),size=0.2)+
    scale_color_viridis_d()

>>>>

```{=tex}
\begin{itemize}
\item  Drop missing latitude and longitude
\item Keep GPS positions indicated by value 1 in \texttt{pos\_org} column
\item  Project to UTM zone 26. The narwhal positions are in latitude and longitude.To analyze the track and visualize it better we project it on a plane in UTM coordinastes.
The positions of the narwhals are in the UTM zone 26 north (in Greenland).
\item Add a column "speed" with empirical horizontal speed obtained from the GPS positions (not interpolated)
\end{itemize}

@@

#remove rows with unknown positions
dataprep=data[!(is.na(data$Lat)) & !(is.na(data$Long)),]

#keep rows indicated as GPS measurements
dataprep=dataprep[dataprep$pos_org==1,]

#number of rows
nrows=length(dataprep[,1])

@@

```{r display land , fig.cap="Land polygons"}
ggplotly(plot_land)
```

@@

```{r compare distances to shore}
plot(res_land$min_distances-dataprep$DistanceShore)
plot(res_coastline$min_distances-dataprep$DistanceShore)
```

```{r summary distances to shore}
summary(res_land$min_distances)
summary(res_coastline$min_distances)
summary(dataprep$DistanceShore)
```
Coastline distances are a bit closer to original distances in the dataset.
Anyway, we shall keep distance computed from the land in the sequel, since the land polygons will be the reference domain
for constrained movement of the narwhals.

@@

check_plot2=ggplot()+geom_sf(data=land_utm$geometry,fill=NA)+
    coord_sf(datum=st_crs(32626))+
  geom_point(data=df,aes(x,y,col=(DistanceShore==0)),size=0.2)+
    scale_color_viridis_d()
